### 2016-08-05 syntax.tex
cheers
changing syntax so that Rep returns public and private aux info

\section{(Set-multiplicity) Data Structures}
\def\pub{{\sf pub}}
\def\aux{{\sf aux}}

@@

\heading{Data Structures.}  We being by defining a general primitive that produces a representation of a given multiset, and provides a mechanism to answer a specified collection of questions about that multiset.  Informally, the representation should serve as a good approximation of the multiset, at least with respect to the allowed questions.
\begin{definition}[Data Structure] \rm
Fix a non-empty set~$\univ$. A \emph{static data structure} (over universe~$\univ$)
is a tuple $\setprim=(\Sigma,\Gamma,\calS,\calQ,\Rep, \Qry)$ where:
\begin{itemize}
\item $\Sigma$ is the \emph{representation alphabet}
\item $\Gamma$ is the set of possible \emph{query answers}
\item $\calS$ is a set  of multisets over universe~$\univ$
\item $\mathcal{Q}=\{q \colon \mathcal{S} \to \Gamma\}$ is a set of supported \emph{queries}
\item $\Rep \colon \mathcal{S} \to \Sigma^* \times \bits^*$ is a randomized \emph{representation algorithm}, taking as input $S \in \mathcal{S}$, and outputting a \emph{representation} $M \in \Sigma^*$ and some \emph{auxilliary information} $\aux \in \bits^*$
\item $\Qry\colon \Sigma^* \times \bits^* \times \calQ \to \Gamma$ is a deterministic \emph{query-evaluation} algorithm, taking as input $M, \aux, q$, and returns \emph{answer} $a \in \Gamma$.
\end{itemize}
\hfill\dqed
\end{definition}
\tsnote{I'm now leaning towards having $\Rep$ output~$M$ and explicit pubic/private auxillary data.  Pushing secrets into~$M$ makes it more difficult to capture (for example) the Neidermeier et al.\ attacks in our privacy notions; see Figure~\ref{fig:privacy}.  In those attacks, the adversary would get the bitarray (which one naturally thinks of as~$M$) but not the secret keys for the PRFs/hash functions.  You could say that the bitarray is put in $\aux$ and that~$M$ \emph{only} contains the secret keys, but this is unnatural.}
All of the data structures that we will consider have the property that~$|M|$ does not depend on the coins of~$\Rep$. Thus we assume that~$|M|$ depends only on~$|S|$.  In general this need not be true, but the assumption does not restrict our syntax from capturing natural constructions (e.g., those seen in practice), and it makes it easier to state our security notions.

\begin{definition}[Error probability and rate] \rm
Fix a data structure $\setprim=(\Sigma,\Gamma,\calS,\calQ,\Rep, \Qry)$.
Let $\mathrm{Err}(S,q,M,\aux)$ be a predicate that is true iff $\Qry(M, \aux, q) \neq q(S)$.  Then a data structure
has \emph{error-probability} $\epsilon$ if
$\max_{S,q}\Pr[(M,\aux) \getsr \Rep(S) : \mathrm{Err}(S,q,M,\aux)=1 ] \leq \epsilon$
where the maximum is over $S \in \mathcal{S}$ and $q \in \mathcal{Q}$.  
%
Relatedly, it has \emph{error-rate} $\tilde{\epsilon}$ if
$\max_{S}\Pr[(M,\aux) \getsr \Rep(S);\, q \getsr \calQ : \mathrm{Err}(S,q,M,\aux)=1] \leq \tilde{\epsilon}$.
\hfill\dqed
\end{definition}

@@

\def\bin{{\sf Bin}}
\heading{Trivial attacks, and error-probability vs.\ error-rate. } 
Let~$\Pi$ be a scheme that is $(1-\epsilon)$-correct, and consider an attack that 
asks random queries $q_1,q_2,\ldots,q_T$, i.e., uniform samples from~$\calQ$.
%The expected number of random queries needed to
%find a error is $1/\epsilon$ tries; thus, an attacker making $O(t/\epsilon)$ queries succeeds with constant probability (almost always).
%Actually, it seems worthwhile to formalize this. 
Let $\mathrm{Err}_i$ be a random variable indicating the event $\Qry(M,\aux,q_i)\neq q_i(S)$.  
The $(1-\epsilon)$-correctness of~$\Pi$ guarantees that the probability that $\mathrm{Err}_i=1$ is at most~$\epsilon$, 
and it is tempting to model the~$\mathrm{Err}_i$ as i.i.d. Bernoulli trials with success probability~$\epsilon$.
However, a moment's reflection shows that this is not the case.  Concretely, consider the following scheme.  Fix $|S|=n$ for all $S \in \calS$, where $n \ll |\univ|$.  For the queries, let $\calQ = \{q_x\colon \calS \to \bits \,|\, \forall x \in \univ\}$ where $\forall S \in \calS$ the predicate $q_x(S)=1 \Leftrightarrow x \in S$.  Let $\Rep(S)$ pick a random element~$s \in S$, and return $M = S \setminus \{s\}$ and $\aux = \emptystring$.  To respond to queries, let~$\Qry(M,\aux,q_x)=1$ iff $x \in M$.  It is easy to see that this scheme is $1/n$-correct; for any fixed~$S$, there are exactly~$n$ queries~$q_x$ that have probability~$1/n$  (over the coins of~$\Rep$) of causing a correctness error, and the remaining $|\univ|-n$ queries have probability zero of causing an error.  Moreover, once the coins of~$\Rep$ are fixed, there is exactly \emph{one} query that will cause an error.  Thus for random queries from~$Q$, for all~$i\in[T]$ we have $\Prob{\mathrm{Err}_i=1}=1/|\univ| \ll 1/n = \epsilon$.

@@

%Our treatment of $M,\aux$ is not symmetric, since we only count $M$ against the size
%of the set-representation. When we define security
%we may give the attacker $\aux$ but not~$M$, and in that case the set-representation cannot
%trivially make everything part of~$\aux$. 
%\tsnote{Encoding secrets into~$M$ may cause some headaches in the security definitions; we'll see. }
%\tsnote{By the way, I'm coming around to the
%  viewpoint that it is fair to include in the ``size'' of the
%  representation any secret information.  In makes more difficult
%  comparisons to classical lower-bounds (which [NY] makes, incorrectly
% IMO), but it does seem fair to say ``this is how many bits of memory
% you need to allocate to represent this set \emph{securely}.''}

%\heading{Bloom Filters and Related Data Structures. }
It is not hard to see Bloom filters are one possible instantiation a set-multiplicity data structure.
Fix an $n \geq 0$, let $\mathcal{S}=[\univ]^n$, and let $\Sigma=\Gamma=\bits$.  
Define the queries so that $q_x(S)=1$ if $x \in S$, and 0 otherwise.  Define algorithm~$\Rep$, with functions $k=k(|S|), m=m(|S|)$ and a description of a hash function family~$\mathcal{H}=\{h \colon \univ \to \{1,2,\ldots,m\}\}$ hardwired into it, to sample hash functions $h_1, \ldots, h_k$ from~$\mathcal{H}$ and then compute an $m$-bit array in the standard way.  It sets $\aux = (h_1, \ldots, h_k)$ and $M$ to be the bit array, and returns these.  The algorithm~$\Qry$ is defined so that, on input $(M,\aux,q_x)$, it returns the minimum of the values (either 0 or 1) held at $M[h_1(x)],\ldots,M[h_k(x)]$.  We note that, while our error-probability and error-rate do not distinguish among ``types'' of errors, they also do not preclude instantiations (like the classical Bloom filter) that admit only one type of error, e.g. false-positives.

As a simple extension, consider $\Sigma=\mathbb{N}$ and a scheme that stores counters at each position in an array.  In particular, for each~$x \in S$ the $\Rep$~algorithm increments the counters stored at positions $h_1(x), h_2(x), \ldots, h_k(x)$, returns the final array of counters as~$M$.  The $\Qry$-algorithm remains the same. Such a construction gives a static version of a counting Bloom filter~\cite{xxx}.  We will capture dynamic versions, as well as count-min sketches~\cite{xxx}, scalable Bloom filters~\cite{xxx}, etc., in a later section.

A Bloom filter with secret hash functions is captured by appending to~$M$ appropriate encodings of $h_1,h_2,\ldots,h_k$, and setting $\aux=\emptystring$.  Related constructions \tsnote{Name them?} that use a secret key~$K$, but not for hashing, are captured by appending~$K$ to~$M$ and setting $\aux=((h_1, \ldots, h_k)$.

>>>>

\section{(Set-multiplicity) Data Structures}

@@

\heading{Data Structures.}  We being by defining a general primitive that produces a representation of a given multiset, and provides a mechanism to answer a specified collection of questions about that multiset.  Informally, the representation should serve as a good approximation of the multiset, at least with respect to the allowed questions.
\begin{definition}[Data Structure] \rm
Fix a non-empty set~$\univ$. A \emph{static data structure} (over universe~$\univ$)
is a tuple $\setprim=(\Sigma,\Gamma,\calS,\calQ,\Rep, \Qry)$ where:
\begin{itemize}
\item $\Sigma$ is the \emph{representation alphabet}
\item $\Gamma$ is the set of possible \emph{query answers}
\item $\calS$ is a set  of multisets over universe~$\univ$
\item $\mathcal{Q}=\{q \colon \mathcal{S} \to \Gamma\}$ is a set of supported \emph{queries}
\item $\Rep \colon \mathcal{S} \to \Sigma^* \times \bits^* \times \bits^*$ is a randomized \emph{representation algorithm}, taking as input $S \in \mathcal{S}$, and outputting a \emph{representation} $M \in \Sigma^*$, \emph{private side-information} $\privaux \in \bits^*$, and \emph{public side-information} $\pubaux \in \bits^*$
\item $\Qry\colon \Sigma^* \times \bits^* \times \calQ \to \Gamma$ is a deterministic \emph{query-evaluation} algorithm, taking as input $M, \privaux, q$, and returns \emph{answer} $a \in \Gamma$.
\end{itemize}
\hfill\dqed
\end{definition}
\tsnote{I implemented this change.}\tsnote{I'm now leaning towards having $\Rep$ output~$M$ and explicit pubic/private auxillary data.  Pushing secrets into~$M$ makes it more difficult to capture (for example) the Neidermeier et al.\ attacks in our privacy notions; see Figure~\ref{fig:privacy}.  In those attacks, the adversary would get the bitarray (which one naturally thinks of as~$M$) but not the secret keys for the PRFs/hash functions.  You could say that the bitarray is put in $\aux$ and that~$M$ \emph{only} contains the secret keys, but this is unnatural.}
All of the data structures that we will consider have the property that~$|M|$ does not depend on the coins of~$\Rep$. Thus we assume that~$|M|$ depends only on~$|S|$.  In general this need not be true, but the assumption does not restrict our syntax from capturing natural constructions (e.g., those seen in practice), and it makes it easier to state our security notions.

\begin{definition}[Error probability and rate] \rm
Fix a data structure $\setprim=(\Sigma,\Gamma,\calS,\calQ,\Rep, \Qry)$.
Let $\mathrm{Err}(S,q,M,\privaux)$ be a predicate that is true iff $\Qry(M, \privaux, q) \neq q(S)$.  Then a data structure
has \emph{error-probability} $\epsilon$ if
$\max_{S,q}\Pr[(M,\pubaux,\privaux) \getsr \Rep(S) : \mathrm{Err}(S,q,M,\privaux)=1 ] \leq \epsilon$
where the maximum is over $S \in \mathcal{S}$ and $q \in \mathcal{Q}$.  
%
Relatedly, it has \emph{error-rate} $\tilde{\epsilon}$ if
$\max_{S}\Pr[(M,\pubaux,\privaux) \getsr \Rep(S);\, q \getsr \calQ : \mathrm{Err}(S,q,M,\privaux)=1] \leq \tilde{\epsilon}$.
\hfill\dqed
\end{definition}

@@

\def\bin{{\sf Bin}}
\heading{Trivial attacks, and error-probability vs.\ error-rate. } 
Let~$\Pi$ be a scheme that is $(1-\epsilon)$-correct, and consider an attack that 
asks random queries $q_1,q_2,\ldots,q_T$, i.e., uniform samples from~$\calQ$.
%The expected number of random queries needed to
%find a error is $1/\epsilon$ tries; thus, an attacker making $O(t/\epsilon)$ queries succeeds with constant probability (almost always).
%Actually, it seems worthwhile to formalize this. 
Let $\mathrm{Err}_i$ be a random variable indicating the event $\Qry(M,\privaux,q_i)\neq q_i(S)$.  
The $(1-\epsilon)$-correctness of~$\Pi$ guarantees that the probability that $\mathrm{Err}_i=1$ is at most~$\epsilon$, 
and it is tempting to model the~$\mathrm{Err}_i$ as i.i.d. Bernoulli trials with success probability~$\epsilon$.
However, a moment's reflection shows that this is not the case.  Concretely, consider the following scheme.  Fix $|S|=n$ for all $S \in \calS$, where $n \ll |\univ|$.  For the queries, let $\calQ = \{q_x\colon \calS \to \bits \,|\, \forall x \in \univ\}$ where $\forall S \in \calS$ the predicate $q_x(S)=1 \Leftrightarrow x \in S$.  Let $\Rep(S)$ pick a random element~$s \in S$, and return $M = S \setminus \{s\}$ and $\privaux = \emptystring$.  To respond to queries, let~$\Qry(M,\privaux,q_x)=1$ iff $x \in M$.  It is easy to see that this scheme is $1/n$-correct; for any fixed~$S$, there are exactly~$n$ queries~$q_x$ that have probability~$1/n$  (over the coins of~$\Rep$) of causing a correctness error, and the remaining $|\univ|-n$ queries have probability zero of causing an error.  Moreover, once the coins of~$\Rep$ are fixed, there is exactly \emph{one} query that will cause an error.  Thus for random queries from~$Q$, for all~$i\in[T]$ we have $\Prob{\mathrm{Err}_i=1}=1/|\univ| \ll 1/n = \epsilon$.

@@

%Our treatment of $M,\privaux$ is not symmetric, since we only count $M$ against the size
%of the set-representation. When we define security
%we may give the attacker $\privaux$ but not~$M$, and in that case the set-representation cannot
%trivially make everything part of~$\privaux$. 
%\tsnote{Encoding secrets into~$M$ may cause some headaches in the security definitions; we'll see. }
%\tsnote{By the way, I'm coming around to the
%  viewpoint that it is fair to include in the ``size'' of the
%  representation any secret information.  In makes more difficult
%  comparisons to classical lower-bounds (which [NY] makes, incorrectly
% IMO), but it does seem fair to say ``this is how many bits of memory
% you need to allocate to represent this set \emph{securely}.''}

%\heading{Bloom Filters and Related Data Structures. }
It is not hard to see Bloom filters are one possible instantiation a set-multiplicity data structure.
Fix an $n \geq 0$, let $\mathcal{S}=[\univ]^n$, and let $\Sigma=\Gamma=\bits$.  
Define the queries so that $q_x(S)=1$ if $x \in S$, and 0 otherwise.  Define algorithm~$\Rep$, with functions $k=k(|S|), m=m(|S|)$ and a description of a hash function family~$\mathcal{H}=\{h \colon \univ \to \{1,2,\ldots,m\}\}$ hardwired into it, to sample hash functions $h_1, \ldots, h_k$ from~$\mathcal{H}$ and then compute an $m$-bit array in the standard way.  It sets $\privaux=\pubaux = (h_1, \ldots, h_k)$ and~$M$ to be the bit array, and returns these.  The algorithm~$\Qry$ is defined so that, on input $(M,\privaux,q_x)$, it returns the minimum of the values (either 0 or 1) held at $M[h_1(x)],\ldots,M[h_k(x)]$.  We note that, while our error-probability and error-rate do not distinguish among ``types'' of errors, they also do not preclude instantiations (like the classical Bloom filter) that admit only one type of error, e.g. false-positives.

As a simple extension, consider $\Sigma=\mathbb{N}$ and modify the basic Bloom filter so that it that stores counters, rather than bits, at each position in an array.  In particular, for each~$x \in S$ the $\Rep$~algorithm increments the counters stored at positions $h_1(x), h_2(x), \ldots, h_k(x)$, returns the final array of counters as~$M$.  The $\Qry$-algorithm remains the same. Such a construction gives a static version of a counting Bloom filter~\cite{xxx}.  We will capture dynamic versions, as well as count-min sketches~\cite{xxx}, scalable Bloom filters~\cite{xxx}, etc., in a later section.

A Bloom filter with secret hash functions is captured by setting $\privaux=(h_1,h_2,\ldots,h_k)$, and setting $\pubaux=\emptystring$.  Related constructions \tsnote{Name them?} that use a secret key~$K$, but not for hashing, are captured setting $\privaux=K, \pubaux=((h_1, \ldots, h_k)$.
